---
title: "02-methods"
output: pdf_document
date: "2025-04-05"
---

```{r}
library(knitr)
library(markdown)
```


```{r}
print("this is a test file")
```

# Methods

This project uses the programming language R to model walksheds of 5 different cities in the United States: Atlanta, Georgia; Houston, Texas; Jacksonville, Florida; Nashville, Tennessee; and Phoenix, Arizona. All of these cities are located in the Sun Belt of the United States, a region characterized by _ and have populations of over 1 million people in their core-based statistical areas. The following steps below will have maps from Phoenix to demonstrate what is being done.

The first step was to map the urban tracts within each city. The urban tracts are created from the city's census tracts, defined as a small, geographic area. The tracts are selected from each city's urban area. Census tracts are good for making comparisons of different areas within a city due to their small size, and that much of the US Census Data is documented for census tracts as well. An example of Phoenix's census tract