---
title: 'My Graduate Thesis'
author: 'You R. Name'
#date: 'May 20xx'
university: 'McMaster University'
faculty: 'Faculty of Science'
department: 'School of Earth, Environment and Society'
subject: 'Geography'
supervisor: 'Antonio Paez'
bdegree: 'B.Sc.' # Bachellor's Degree
mdegree: ''   # Master's Degree (if any)
# This is the degree you are aiming for with this thesis
degree: 'Doctor of Philosophy'

knit: bookdown::render_book
site: bookdown::bookdown_site

output:
  macdown::thesis_pdf: default 
  keep_tex: true

lay_abstract: |
  The lay abstract must be 150 words or less.
  
  It must explain the key goals and contributions of the thesis in lay terms that are accessible to the general public. 
abstract: |
  This is the abstract.
  
  I can write a really long abstract.
# Note that a tab is needed on the line after the `|`. Subsequent paragraphs need to be indented too.
acknowledgements: |
  I want to thank a few people.
  
  This includes my friends.
dedication: |
  You can have a dedication here if you wish. 
declarations: |
  I did most of the research.
  
  Also the writting. 
  
  Sometimes I cried.
  
  But mostly I had fun.
quotation: 'You may hate gravity, but gravity does not care'
# Author of the quotation
qauthor: 'Albert Einstein'
# Specify the location where the bibliography file is found below
bibliography: bib/thesis.bib
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
---

# This is the degree you are aiming for with this thesis

Placeholder



<!--chapter:end:index.Rmd-->

---
title: "00-thesis-draft"
output: pdf_document
date: "2025-04-05"
---

```{r}
if (!require("remotes")) 
  install.packages("remotes", 
                   repos = "https://cran.rstudio.org")
if (!require("bookdown")) 
  install.packages("bookdown", 
                   repos = "https://cran.rstudio.org")
remotes::install_github("paezha/macdown")
```


<!--chapter:end:00-thesis-draft.Rmd-->


# R Markdown Basics {#rmd-basics}

Placeholder



<!--chapter:end:01-chap1.Rmd-->


# Mathematics and Science {#math-sci}

Placeholder


## Math
### Other examples of reactions
## Physics
## Biology

<!--chapter:end:02-chap2.Rmd-->

---
title: "02-methods"
output: pdf_document
date: "2025-04-05"
---

```{r}
library(tidycensus)
library(tigris)
library(dplyr) # A Grammar of Data Manipulation
library(ggraph) # An Implementation of Grammar of Graphics for Graphs and Networks
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(glue) # Interpreted String Literals
library(here) # A Simpler Way to Find Your Files
library(igraph) # Network Analysis and Visualization
library(sf) # Simple Features for R
# library(sfnetworks) # Tidy Geospatial Networks
library(tidygraph) # A Tidy API for Graph Manipulation
library(tidyr) # Tidy Messy Data
# library(r5r) # Rapid Realistic Routing with 'R5'
# library(h3) # R Bindings for H3 # R Bindings for H3
library(purrr) # Functional Programming Tools
library(units) # Measurement Units for R Vectors
library(ggpubr)
```


```{r}
print("this is a test file")
```

```{r, echo=FALSE}
# Walksheds
load(file = glue::glue(here::here(),
                       "/data/houston_walksheds.rda"))

# Tracts
load(file = glue::glue(here::here(),
                       "/data/houston_tracts.rda"))

# Urban boundaries
load(file = glue::glue(here::here(),
                       "/data/houston_urban.rda"))

# CBSA boundary
load(file = glue::glue(here::here(),
                       "/data/houston_cbsa.rda"))

# Houston dataframe
load(file = glue::glue(here::here(),
                       "/data/houston_tracts_df.rda"))

# Complete dataframe
load(file = glue::glue(here::here(),
                       "/data/combined_df.rda"))
```


# Methods

This project uses the programming language R to model walksheds of 5 different cities in the United States: Atlanta, Georgia; Houston, Texas; Jacksonville, Florida; Nashville, Tennessee; and Phoenix, Arizona. All of these cities are located in the Sun Belt of the United States, a region characterized by _ and have populations of over 1 million people in their core-based statistical areas. The following steps below will have maps from Houston to demonstrate what is being done.

The first step was to map the urban tracts within each city. The urban tracts are created from the city's census tracts, defined as a small, geographic area. The tracts are selected from each city's urban area. Census tracts are good for making comparisons of different areas within a city due to their small size, and that much of the US Census Data is documented for census tracts as well. Atlanta contains 1390 tracts, Houston contains 1376 tracts, Jacksonville contains 281 tracts, Nashville contains 287 tracts, Phoenix contains 981 tracts. An example of Phoenix's census tract map can be seen below, with the core-based statistical area shown in grey, and its urban tracts outlined in blue.

```{r, echo=FALSE}
ggplot() +
  geom_sf(data = houston_cbsa,
          fill = "lightgray") +
  geom_sf(data = houston_tracts,
          color = "blue",
          fill = NA) +
  theme_bw()
```

Next, walksheds for each census tract were generated. These walksheds determine how much a person can travel from a given starting point within 15 minutes. In this case, based on the r5r package used for generating these walksheds, the theoretical maximum a person could travel within 15 minutes walking at the base speed of _ and no obstacles would be 2544690 meters squared. Hence, each walkshed would have an area between 0 and 2544690 meters squared. They were generated for the centroids of each census tract. An example walkshed can be seen below in blue, with the census tract being outlined in grey.

```{r, echo=FALSE}
test_unit = 98
#98

ggplot() +
  geom_sf(data = houston_tracts |>
            filter(GEOID == houston_walksheds$GEOID[test_unit]),
          color = "black",
          fill= "lightgray") + 
  geom_sf(data = houston_walksheds[test_unit,],
          color = "blue",
          fill = NA,
          linewidth = 1) +
  theme_bw()
```

The shape and area of the walksheds depend on the street patterns present, and obstacles present such as roads.

Afterwards, the US Census Data for each city and its tracts is collected using the Tidycensus package. The main demographics used include average household income (as a control variable), white population count, black population count, and other _, each being at a census tract level.

Afterwards, proportion variables are calculated. This includes the white population proportion (out of 1.00) and the black population proportion (out of 1.00). These are calculated by dividing white and black population counts by total population counts (calculated by adding _) respectfully. The results can then be mapped to view, for instance, black population proportions per census tract in each city, as seen in the example Houston map below:

```{r, echo=FALSE}
ggplot(houston_tracts_df, aes(fill = black_prop)) + 
  geom_sf() +
  theme_void() +
  scale_fill_viridis_c(option = "turbo") +
  labs(fill = "Proportion of
Black Population")
```

Or, to map the walkshed areas (in meters squared) for each census tracts of the cities:

```{r, echo=FALSE}
# Make sure Census Data covers all of the tracts
ggplot(houston_tracts_df, aes(fill = area)) + 
  geom_sf() + 
  theme_void() +
  scale_fill_viridis_c(option = "turbo") +
  labs(fill = "Area of Walkshed
in Square Meters")
```

However, maps alone are insufficient in comparing the correlation between walkshed areas and black populations of each census tract. Hence, scatter plots are created to compare the two for each city. Below is an example of Houston's black population proportion (x-axis) compared to the walkshed area of each census tract in meters squared (y-axis):

```{r}
ggscatter(houston_tracts_df, x = "black_prop", y = "area", add = "reg.line", shape = 19, size = 1, xlab = "Proportion of Black Population in Census Tract", ylab = "Area of Walkshed in Square Meters", font.label = c(5, "plain", "black")) +
  stat_cor(label.x = 0.75, label.y = 2000000) #label placement
```

The process was repeated to graph the proportion of white populations and average household income to the walkshed areas in square meters:

```{r}
ggscatter(houston_tracts_df, x = "white_prop", y = "area", add = "reg.line", shape = 19, size = 1, xlab = "Proportion of White Population in Census Tract", ylab = "Area of Walkshed in Square Meters", font.label = c(5, "plain", "black")) +
  stat_cor(label.x = 0.75, label.y = 2000000) #label placement
```

```{r}
ggscatter(houston_tracts_df, x = "income", y = "area", add = "reg.line", shape = 19, size = 1, xlab = "Average Income in Census Tract (USD)", ylab = "Area of Walkshed in Square Meters", font.label = c(5, "plain", "black")) +
  stat_cor(label.x = 110000, label.y = 2000000) #label placement
```



<!--chapter:end:02-methods.Rmd-->

# Conclusion {-}

If we don't want Conclusion to have a chapter number next to it, we can add the `{-}` attribute.

**More info**

And here's some other random info: the first paragraph after a chapter title or section head _shouldn't be_ indented, because indents are to tell the reader that you're starting a new paragraph. Since that's obvious after a chapter or section title, proper typesetting doesn't add an indent there.


<!--chapter:end:03-conclusion.Rmd-->


# References {-}

Placeholder



<!--chapter:end:99-references.Rmd-->

