---
title: "1940 Data"
---

# 1940 Data

This will be the basic data pre-processing work for Phoenix, Arizona. The other city notebooks follow the same steps but with different sets of CBSA data.

Estimated time to run: a few minutes

Load packages:

```{r}
library(readr) # Read Rectangular Text Data
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(glue) # Interpreted String Literals
library(here) # A Simpler Way to Find Your Files
library(sf) # Simple Features for R
library(tidycensus) # Load US Census Boundary and Attribute Data as 'tidyverse' and 'sf'-Ready Data Frames
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(tigris) # Load Census TIGER/Line Shapefiles
```

```{r}
us_1940_demo <- read_csv(glue(here::here(),
                               "/data-raw/1940-data/nhgis0001_csv/nhgis0001_ds76_1940_tract.csv"))

us_1940_tract <- st_read(glue(here::here(),
                               "/data-raw/1940-data/nhgis0001_shape/nhgis0001_shapefile_tl2000_us_tract_1940/US_tract_1940.shp"))
```

Check whether the geometry of 1940 tracts is valid:
```{r}
sum(st_is_valid(us_1940_tract))
```



```{r load files part 1}
# Tracts
load(file = glue::glue(here::here(),
                       "/data/houston_tracts.rda"))

# Urban boundaries
load(file = glue::glue(here::here(),
                       "/data/houston_urban.rda"))

# CBSA boundary
load(file = glue::glue(here::here(),
                       "/data/houston_cbsa.rda"))

# Walksheds
load(file = glue::glue(here::here(),
                       "/data/houston_walksheds.rda"))
```

There is at least one invalid geometry. Fix:
```{r}
us_1940_tract <- us_1940_tract |>
  st_make_valid()
```

Read urban boundaries and tracts:
```{r}
load(glue::glue(here::here(), "/data/houston_cbsa.rda"))
load(glue::glue(here::here(), "/data/houston_tracts.rda"))
```

Check the crs:
```{r}
st_crs(us_1940_tract)
st_crs(houston_cbsa)
```

Transform the 1940 tracts to the same crs as Houston:
```{r}
us_1940_tract <- us_1940_tract |>
  st_transform(crs = st_crs(houston_cbsa))
```

Check that the crs match:
```{r}
st_crs(us_1940_tract) == st_crs(houston_cbsa)
```

Filter 1940 tracts based on Houston CBSA:
```{r}
houston_1940_tracts <- us_1940_tract |>
  st_filter(houston_cbsa)
```

Plot:
```{r}
ggplot() + 
  geom_sf(data = houston_1940_tracts,
          color = "blue",
          fill = NA,
          linewidth = 1) +
  geom_sf(data = houston_tracts,
          color = "red",
          fill = NA,
          linewidth = 0.1)
```

