---
title: "Houston Texas Ped Sheds"
---

A [ped shed](https://pedshed.net/?page_id=5), also called a pedestrian shed, is the measure of how far a person can walk starting at a given location within a given time range. This is heavily dependent on street layouts, which can impact how long it would take a person to reach \_. This term will be used interchangeably with "walksheds" throughout this notebook.

Load packages:

```{r}
options(java.parameters = "-Xmx30G")
library(dplyr) # A Grammar of Data Manipulation
library(ggraph) # An Implementation of Grammar of Graphics for Graphs and Networks
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(glue) # Interpreted String Literals
library(here) # A Simpler Way to Find Your Files
library(igraph) # Network Analysis and Visualization
library(sf) # Simple Features for R
library(sfnetworks) # Tidy Geospatial Networks
library(tidygraph) # A Tidy API for Graph Manipulation
library(tidyr) # Tidy Messy Data
library(r5r) # Rapid Realistic Routing with 'R5'
library(h3) # R Bindings for H3 # R Bindings for H3
library(purrr) # Functional Programming Tools
library(units) # Measurement Units for R Vectors
```

## Setting up the Data:

Load data:
```{r load files part 1}
# Tracts
load(file = glue::glue(here::here(),
                       "/data/houston_tracts.rda"))

# Urban boundaries
load(file = glue::glue(here::here(),
                       "/data/houston_urban.rda"))

# CBSA boundary
load(file = glue::glue(here::here(),
                       "/data/houston_cbsa.rda"))
```

Bounding box of CBSA:
```{r}
st_bbox(houston_cbsa)
```

## Network data

Retrieve OSM data from [BBBike](https://extract.bbbike.org) using the coordinates of the bounding box. It will be a proto-buffer file (pbf). Save under `data-raw/r5_graph/$CITY_NAME$`.

```{r set up r5 path}
r5_path <- file.path(glue::glue(here::here(), "/data-raw/r5_graph/Houston/"))
```

Build the graph:
```{r build-graph}
r5_cma <- setup_r5(data_path = r5_path, 
                            verbose = FALSE)

# requires JDK version 21 (exactly version 21) to run
```

The origins are coordinates of centroids tracts and the destinations the coordinates of the h3 geohashes. 

Obtain the tract centroids (the geometry is degenerate and needs to be fixed):
```{r prepare-inputs}
origins_da_houston <- houston_tracts |>
  st_make_valid() |>
  st_centroid() |>
  # Transform to lat-long
  st_transform(crs = st_crs(4326))
```

Organize the data for the origins to match the format expected by {r5r} (id, lon, lat):
```{r}
origins_i_tract <- data.frame(id = houston_tracts$GEOID,
                           origins_da_houston |>
                          st_coordinates()) |>
  rename(lon = X, lat = Y) |>
  dplyr::select(id, lon, lat)
```

This is the maximum area with a walking speed of 3.6 km/h and a time of 15 min:
```{r}
pi * (units::set_units(3.6, km/h) |> set_units(m/min) * set_units(15, min))^2
```

Obtain an h3 that covers a larger area:
```{r}
h3_i <- geo_to_h3(c(origins_i_tract$lat[1], origins_i_tract$lon[1]), res = 8) |>
  k_ring(1) |> h3_set_to_multi_polygon()

h3_i |> st_area()
```

Obtain the children of the chosen polygon(s):
```{r}
i <- 5
h3_i <- geo_to_h3(c(origins_i_tract$lat[i], origins_i_tract$lon[i]), res = 7) |>
  k_ring(1)

h3_j <- purrr::map(h3_i,
                   .f = ~ h3_to_children(.x,
                                  res = 13)) |>
  unlist()
```

Check coverage of the children (make sure that the parents are fully covered):
```{r}
ggplot() + geom_sf(data = h3_i |> h3_set_to_multi_polygon()) + 
  geom_sf(data = h3_to_geo_boundary_sf(h3_j), fill = NA)
```

Obtain the coordinates of the h3 destinations:
```{r set destinations}
destinations_h3 <- h3_to_geo(h3_j)
```

Likewise, organize the data for the origins to match the format expected by {r5r} (id, lon, lat):
```{r set destination sites}
destinations_j_h3 <- data.frame(id = h3_j, 
                             lat = destinations_h3[,1],
                             lon = destinations_h3[,2])
```

Test routing with a single origin:
```{r}
ttm_i <- travel_time_matrix(r5_cma,
                                  origins = origins_i_tract[i,],
                                  destinations = destinations_j_h3,
                                  mode = "WALK",
                                  departure_datetime = as.POSIXct(strptime("2021-04-05 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                                  max_walk_time = 15, # minutes
                                  max_trip_duration = 15)
```

Retrieve the walkshed as a multipolygon:
```{r}
walksheds_i <- h3_set_to_multi_polygon(ttm_i$to_id) |> 
  as_tibble() |> st_as_sf(crs = 4326)
```

Compare the walkshed to the parent h3s:
```{r}
ggplot() +
  geom_sf(data = h3_to_geo_boundary_sf(h3_i), 
          fill = NA) +
  # geom_sf(data = h3_to_geo_boundary_sf(h3_j), 
  #         fill = "blue") +
  geom_sf(data = walksheds_i)
```

Now implement for the rest of origins:
```{r}
n_origins <- nrow(origins_i_tract)

start.time <- Sys.time()

for(i in 1:n_origins){
  
  # Obtain the h3 polygons around the centroid of the tract at resolution 7
  h3_i <- geo_to_h3(c(origins_i_tract$lat[i], origins_i_tract$lon[i]), res = 7) |>
    k_ring(1)
  
  # Obtain the children of the res 7 polygons at resolution 13, whith are about 50 square meters in size
  h3_j <- purrr::map(h3_i,
                     .f = ~ h3_to_children(.x,
                                           res = 13)) |>
    unlist()
  
  # Obtain the coordinates of the res 13 polygons to use as destinations for origin i
  destinations_h3 <- h3_to_geo(h3_j)
  
  # Organize the data for the origins to match the format expected by {r5r} (id, lon, lat)
  destinations_j_h3 <- data.frame(id = h3_j, 
                                  lat = destinations_h3[,1],
                                  lon = destinations_h3[,2])
  
  # Routing for origin i
  ttm_i <- bind_rows(ttm_i,travel_time_matrix(r5_cma,
                                              origins = origins_i_tract[i,],
                                              destinations = destinations_j_h3,
                                              mode = "WALK",
                                              departure_datetime = as.POSIXct(strptime("2021-04-05 08:00:00", "%Y-%m-%d %H:%M:%S", tz = "EST5EDT")),
                                              max_walk_time = 15, # minutes
                                              max_trip_duration = 15))
}

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

The above chunk took 1.37 hours in Antonio's system Rogue.

Retrieve the h3 areas as multipolygons:
```{r}
start.time <- Sys.time()

houston_walksheds <- ttm_i |> 
  group_by(from_id) |>
  group_map(~ h3_set_to_multi_polygon(.x$to_id),
            .keep = TRUE)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

The above chunk took about 1.18 h in Antonio's system Rogue.

Obtain the GEOIDs of the walksheds:
```{r}
houston_walkshed_IDs <- ttm_i |>
  group_by(from_id) |>
  summarize(from_id = first(from_id),
            .groups = "drop")
```

Convert to simple features identified by the origin:
```{r}
start.time <- Sys.time()

houston_walksheds <- data.frame(GEOID = houston_walkshed_IDs$from_id,
                         geometry = purrr::map_df(houston_walksheds,
                            tibble::as_tibble)) |>
  st_as_sf(crs = 4326)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

Verify the catchments:
```{r}
test_unit = sample.int(nrow(houston_walksheds), 1)

ggplot() +
  geom_sf(data = houston_tracts |>
            filter(GEOID == houston_walksheds$GEOID[test_unit]),
          color = "black",
          fill= "lightgray") + 
  geom_sf(data = houston_walksheds[test_unit,],
          color = "blue",
          fill = NA,
          linewidth = 1) #+
  #geom_sf(data = roads)
```

Save walksheds:
```{r save origin and destination files}
# Origins
save(houston_walksheds, 
     file = glue::glue(here::here(),
                       "/data/houston_walksheds.rda"),
     compress = "bzip2")

```